% Data

## Load libraries

```{r}
library(readxl)
library(tidyverse)
library(lubridate)
```

## Marsden Farm

### Canopy reflectance

#### Read data

- The raw data was organized in different Excel sheets based on study year. 
- I'll upload the data using `read_excel` and `read_csv` functions and will run some cleaning before merging all data sets. 

```{r warning=FALSE}
MR2016 <- read_excel("data/Marsden RFL 2016.xlsx")
MR2017 <- read_excel("data/Marsden RFL 2017.xlsx")
MR2018 <- read_csv("data/Marsden RFL 2018.csv", 
                   col_types = cols(DATE = col_date(format = "%m/%d/%Y")))
```

#### Data cleaning

- In each marked soybean quadrat (observational unit), we took two readings of remote sensing. 
- In 2018 data set, the mean reflectance value (single reflectance value/quadrat) was not calculated like other year's data. 
- Thus, I'll calculate the mean reflectance and then will gather the data using `gather` function. 

```{r message=FALSE, warning=FALSE}
MR2018 <- MR2018 %>%
  group_by(DATE, Block, Plot, Rotation, Quadrat) %>%
  summarise(`460` = mean(W460, na.rm = TRUE),`510` = mean(W510, na.rm = TRUE),
            `560` = mean(W560, na.rm = TRUE),`610` = mean(W610, na.rm = TRUE),
            `660` = mean(W660, na.rm = TRUE),`710` = mean(W710, na.rm = TRUE),
            `760` = mean(W760, na.rm = TRUE),`810` = mean(W810, na.rm = TRUE)) %>%
  gather(Wavelength, Reflectance, `460`:`810`, na.rm = TRUE, convert = TRUE)
```

#### Combining all data sets

- Now, I'll combine all of the reflectance data sets from Marsden Farm and will save the final file as CSV. 

```{r}
# Cmobine all data sets
Marsden_RFL <- rbind(MR2016, MR2017, MR2018)

# Store data as .csv file
write.csv(Marsden_RFL, "data/Marsden_RFL.csv")
```

### SDS severity and incidence

#### Read data

```{r}
Marsden_SDS <- read_csv("data/Marsden_SDS.csv")
```


## Hinds Farm

- The data from Hinds Farm is of only one year. 

### Read data

```{r}
HR2018 <- read_csv("data/Hinds RFL 2018.csv", 
                   col_types = cols(DATE = col_date(format = "%m/%d/%Y")))

```

### Data cleaning

- Hinds data needs some cleaning like renaming some columns, adding original variety names, calculating mean canopy reflectance for each plot and gathering the canopy reflectance data. 

```{r message=FALSE}
HR2018 <- HR2018 %>%
  mutate(Variety = ifelse(Variety  == "1", "Pioneer P20T79R2", "Syngenta NK Soybean S24-K2"))%>%
  separate(Reading, into = c("Plot1", "Plot2"), sep = c(3), remove = FALSE, convert = TRUE) %>%
  select(-Plot, -Plot2) %>%
  rename(Plot = Plot1, Concentration = Conc) %>%
  group_by(DATE, Block, Plot, Treatment, Variety, Concentration) %>%
  summarise(`460` = mean(W460, na.rm = TRUE),`510` = mean(W510, na.rm = TRUE),
            `560` = mean(W560, na.rm = TRUE),`610` = mean(W610, na.rm = TRUE),
            `660` = mean(W660, na.rm = TRUE),`710` = mean(W710, na.rm = TRUE),
            `760` = mean(W760, na.rm = TRUE),`810` = mean(W810, na.rm = TRUE)) %>%
  gather(Wavelength, Reflectance, `460`:`810`, na.rm = TRUE, convert = TRUE)
```

### Store data as .csv file

```{r}
write.csv(HR2018, "data/Hinds_RFL.csv")
```

## Join Reflectance and Disease data

### Marsden Farm 2016

```{r}
# Read disease data
Marsden_SDS <- read_csv("data/Marsden_SDS.csv")

# Data cleaning
MS_2016 <- Marsden_SDS %>%
  filter(ID == "SDS_2016") %>%
  gather(key, value, Severity, Incidence, na.rm = TRUE, convert = TRUE) %>%
  unite(Date_Severity, Date, key, sep = "_", remove = FALSE, na.rm = FALSE) %>%
  select(-Date, -key) %>%
  spread(Date_Severity, value) %>%
  select(-Block, -Plot, -Rotation)
  
# Joining the data sets
MRS_2016 <- Marsden_RFL %>% 
  filter(year(DATE) == 2016) %>%
  mutate(Quadrat = as.numeric(Quadrat)) %>% 
  left_join(MS_2016, by = c("Quadrat" = "Quadrat"))
```

### Store data as .csv file

```{r}
write.csv(MRS_2016, "data/MRS_2016.csv")
```

### Marsden Farm 2017

```{r}
# Data cleaning
MS_2017 <- Marsden_SDS %>%
  filter(ID == "SDS_2017") %>%
  gather(key, value, Severity, Incidence, na.rm = TRUE, convert = TRUE) %>%
  unite(Date_Severity, Date, key, sep = "_", remove = FALSE, na.rm = FALSE) %>%
  select(-Date, -key) %>%
  spread(Date_Severity, value) %>%
  select(-Block, -Plot, -Rotation)
  
# Joining the data sets
MRS_2017 <- Marsden_RFL %>% 
  filter(year(DATE) == 2017) %>%
  mutate(Quadrat = as.numeric(Quadrat)) %>% 
  left_join(MS_2017, by = c("Quadrat" = "Quadrat")) 
```

### Store data as .csv file

```{r}
write.csv(MRS_2017, "data/MRS_2017.csv")
```

### Marsden Farm 2018

```{r}
# Data cleaning
MS_2018 <- Marsden_SDS %>%
  filter(ID == "SDS_2018") %>%
  gather(key, value, Severity, Incidence, na.rm = TRUE, convert = TRUE) %>%
  unite(Date_Severity, Date, key, sep = "_", remove = FALSE, na.rm = FALSE) %>%
  select(-Date, -key) %>%
  spread(Date_Severity, value) %>%
  select(-Block, -Plot, -Rotation)
  
# Not in multiple function
`%nin%` = Negate(`%in%`)

# Joining the data sets
MRS_2018 <- Marsden_RFL %>% 
  filter(year(DATE) == 2018 & Quadrat %nin% c(4811, 4812, 4911, 4912, 4913)) %>%
  mutate(Quadrat = as.numeric(Quadrat)) %>%
  left_join(MS_2018, by = c("Quadrat" = "Quadrat")) %>%
  filter(!is.na(ID))
```

### Store data as .csv file

```{r}
write.csv(MRS_2018, "data/MRS_2018.csv")
```

### Hinds Farm 2018

```{r}
#  Read disease data
Hinds_SDS <- read_csv("data/Hinds_SDS.csv")


# Data cleaning
HS_2018 <- Hinds_SDS %>%
  gather(key, value, Severity, Incidence, na.rm = TRUE, convert = TRUE) %>%
  unite(Date_Severity, Doy, key, sep = "_", remove = FALSE, na.rm = FALSE) %>%
  select(-Doy, -key) %>%
  spread(Date_Severity, value) %>%
  select(-Block, -Treatment, -Variety, -Concentration)

# Joining the data sets
HRS_2018 <- HR2018 %>% 
  left_join(HS_2018, by = c("Plot" = "Plot")) 
```

### Store data as .csv file

```{r}
write.csv(HRS_2018, "data/HRS_2018.csv")
```
